```{r fig-distribution, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Value distribution of the mental health and quality of life scores, stress score and physical performance loss.'}

suppl_figures$distribution$plot

```

__Figure \@ref(fig:fig-distribution). Value distribution of the mental health and quality of life scores, stress score and physical performance loss.__



```{r fig-obs-time, fig.width = 7.086614166, fig.height = 9.055118101, fig.cap = 'Effect of the observation time on mental health and quality of life scoring.'}

suppl_figures$obs_time$plot

```

__Figure \@ref(fig:fig-obs-time). Effect of the observation time on mental health and quality of life scoring.__



```{r fig-survey, fig.width = 7.086614166, fig.height = 9.055118101, fig.cap = 'Effect of the survey duration on mental health and quality of life scoring.'}

suppl_figures$survey$plot

```

__Figure \@ref(fig:fig-survey). Effect of the survey duration on mental health and quality of life scoring.__



```{r fig-correlation-responses, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Inter-correlation of the mental health and quality of life score variables.'}

suppl_figures$correlation_responses$plot

```

__Figure \@ref(fig:fig-correlation-responses). Inter-correlation of the mental health and quality of life score variables.__



```{r fig-correlation-responses-da, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Inter-correlation of the mental health and quality of life score variables in participants with pre-existing depression or anxiety.'}

suppl_figures$correlation_responses_da$plot

```

__Figure \@ref(fig:fig-correlation-responses-da). Inter-correlation of the mental health and quality of life score variables in participants with pre-existing depression or anxiety.__



```{r fig-rforest-anx, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Random Forest modeling of the anxiety scoring.'}

suppl_figures$rforest_anx$plot

```

__Figure \@ref(fig:fig-rforest-anx). Random Forest modeling of the anxiety scoring.__



```{r fig-rforest-dpr, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Random Forest modeling of the depression scoring.'}

suppl_figures$rforest_dpr$plot

```

__Figure \@ref(fig:fig-rforest-dpr). Random Forest modeling of the depression scoring.__



```{r fig-rforest-omh, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Random Forest modeling of the overall mental health scoring.'}

suppl_figures$rforest_omh$plot

```

__Figure \@ref(fig:fig-rforest-omh). Random Forest modeling of the overall mental health scoring.__



```{r fig-rforest-qol, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Random Forest modeling of the quality of life scoring.'}

suppl_figures$rforest_qol$plot

```

__Figure \@ref(fig:fig-rforest-qol). Random Forest modeling of the quality of life scoring.__



```{r fig-perc-explained, fig.width = 7.086614166, fig.height = 7.87401574, fig.cap = 'Fraction of mental health and quality of life scoring deviance explained by the common most influential factors.'}

suppl_figures$perc_explained$plot

```

__Figure \@ref(fig:fig-perc-explained). Fraction of mental health and quality of life scoring deviance explained by the common most influential factors.__



```{r fig-rf-performance-da, fig.width = 7.086614166, fig.height = 3.149606296, fig.cap = 'Prediction of mental health and quality of life scoring by Random modeling in participants with pre-existing depression or anxiety.'}

suppl_figures$rf_performance_da$plot

```

__Figure \@ref(fig:fig-rf-performance-da). Prediction of mental health and quality of life scoring by Random modeling in participants with pre-existing depression or anxiety.__



```{r fig-clust-qc, fig.width = 7.086614166, fig.height = 6.299212592, fig.cap = 'Development of the participant clusters in respect to the common most influential factors for mental health and quality of life scoring.'}

suppl_figures$clust_qc$plot

```

__Figure \@ref(fig:fig-clust-qc). Development of the participant clusters in respect to the common most influential factors for mental health and quality of life scoring.__



```{r fig-clust-features-summary, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Characteristic of the mental disorder risk clusters.'}

suppl_figures$clust_features_summary$plot

```

__Figure \@ref(fig:fig-clust-features-summary). Characteristic of the mental disorder risk clusters.__



```{r fig-clust-features-heat, fig.width = 7.086614166, fig.height = 6.299212592, fig.cap = 'Variables significantly differing between the mental disorder risk clusters.'}

suppl_figures$clust_features_heat$plot

```

__Figure \@ref(fig:fig-clust-features-heat). Variables significantly differing between the mental disorder risk clusters.__

