---
title: "Who is at risk of poor mental health following COVID-19 outpatient management?"
subtitle: "Figures and Tables"
author: "Health after COVID-19 in Tyrol study team"
date: "`r format(Sys.time(), '%Y-%m-%d')`"

output: 
  pdf_document:
    latex_engine: xelatex
bibliography: cov_biblio.bib
csl: american-medical-association.csl
header-includes:
  \usepackage{longtable}
  \usepackage{tabu}
  \usepackage{caption}
  \usepackage{makecell}
  \usepackage{pdflscape}
  \usepackage{array}
  \usepackage{booktabs}
  \usepackage{threeparttable}
  \usepackage{threeparttablex}
  \usepackage{wrapfig}
  \usepackage{multirow}
  \usepackage[normalem]{ulem}
  \usepackage{colortbl}
  \usepackage{xcolor}
  \usepackage{float} \floatplacement{figure}{H} \floatplacement{table}{H}
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}}       \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
---

```{r, setup, include = FALSE}

library(bookdown)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, dev = "cairo_pdf")


```

\newpage

# Tables {#sec:tbl}

```{r tab-baseline-features}

paper_tables$cohort_features$baseline_vars %>% 
  format_line %>%  
  kable(caption = '\\textbf{Baseline characteristic of the study cohorts.}\\\\AT: Austria/Tyrol cohort, IT: Italy/South Tyrol cohort, Test: statistical test used for the AT vs IT comparison, Significance: test p value corrected for multiple comparisons with Benjamini-Hochberg method.', 
        format = 'latex', 
        escape = F, 
        booktabs = T, 
        longtable = T) %>% 
  column_spec(1, width = '3.3cm') %>% 
  column_spec(2:3, width = '4cm') %>% 
  column_spec(4, width = '0.6cm') %>% 
  column_spec(5, width = '2cm') %>% 
  collapse_rows(columns = 1) %>% 
  kable_styling(full_width = F, 
                latex_options = c('striped', 
                                  'hold_position', 
                                  'repeat_header'))

```

\newpage


```{r tab-course-features}

paper_tables$cohort_features$course_vars %>% 
  format_line %>%  
  kable(caption = '\\textbf{Characteristic of the course of SARS-CoV2 infection and convalescence in the study cohorts.}\\\\AT: Austria/Tyrol cohort, IT: Italy/South Tyrol cohort, Test: statistical test used for the AT vs IT comparison, Significance: test p value corrected for multiple comparisons with Benjamini-Hochberg method.', 
        format = 'latex', 
        escape = F, 
        booktabs = T, 
        longtable = T) %>% 
  column_spec(1, width = '3.3cm') %>% 
  column_spec(2:3, width = '4cm') %>% 
  column_spec(4, width = '0.6cm') %>% 
  column_spec(5, width = '2cm') %>% 
  collapse_rows(columns = 1) %>% 
  kable_styling(full_width = F, 
                latex_options = c('striped', 
                                  'hold_position', 
                                  'repeat_header'))

```

\newpage

```{r tab-psych-features}

paper_tables$cohort_features$psych_vars %>% 
  format_line %>%  
  kable(caption = '\\textbf{Rating of the mental health following COVID-19 in the study cohorts.}\\\\AT: Austria/Tyrol cohort, IT: Italy/South Tyrol cohort, Test: statistical test used for the AT vs IT comparison, Significance: test p value corrected for multiple comparisons with Benjamini-Hochberg method.', 
        format = 'latex', 
        escape = F, 
        booktabs = T, 
        longtable = T) %>% 
  column_spec(1, width = '3.3cm') %>% 
  column_spec(2:3, width = '4cm') %>% 
  column_spec(4, width = '0.6cm') %>% 
  column_spec(5, width = '2cm') %>% 
  collapse_rows(columns = 1) %>% 
  kable_styling(full_width = F, 
                latex_options = c('striped', 
                                  'hold_position', 
                                  'repeat_header'))

```


\newpage

# Figures {#sec:figures}

```{r fig-consort, fig.width = mm_inch(paper_figures$consort$w), fig.height = mm_inch(paper_figures$consort$h), dpi = 600, fig.cap = 'CONSORT flow diagram for the study cohorts.'}

paper_figures$consort$plot

```

__Figure \@ref(fig:fig-consort). CONSORT flow diagram for the study cohorts.__

\newpage

```{r fig-rf-components, fig.width = mm_inch(paper_figures$rf_components$w), fig.height = mm_inch(paper_figures$rf_components$h), dpi = 600, fig.cap = 'The most influential factors for the combined mental health and quality of life scoring following COVID-19.'}

paper_figures$rf_components$plot

```

__Figure \@ref(fig:fig-rf-components). The most influential factors for the combined mental health and quality of life scoring following COVID-19.__

The effects of `r length(rforest$variables)` demographic, clinical, socioeconomic and psychosocial factors (__Supplementary Table S1__) on the overall mental health, quality of life, anxiety and depression scoring were modeled with random forest technique (__Supplementary Figures S3 - S6__). The the impacts of candidate factors on the model fits were subjected to two-dimensional centered principal component analysis. Factors' loadings in respect to the principal components (PC) are presented in the plots, percent variance is shown in the axes. Top 10 factors with the largest loadings vectors as a measure of the influence on the combined mental health and quality of life scoring were labeled, point color corresponds to the vector length.

prol.: prolonged, SIF: severe illness feeling, imp.: impaired, conc.: concentration, #: number, tired.day.: tiredness at day, pre-cov sleep disord.: sleep disorder before COVID-19, 3Q, 4Q: 3^rd^ and 4^th^ quartile, persist.: persistent.

\newpage

```{r fig-uni-modeling, fig.width = mm_inch(paper_figures$uni_modeling$w), fig.height = mm_inch(paper_figures$uni_modeling$h), dpi = 600, fig.cap = 'Correlation of the most influential factors with the mental health readouts investigated by univariable modeling.'}

paper_figures$uni_modeling$plot

```

__Figure \@ref(fig:fig-uni-modeling). Correlation of the most influential factors with the mental health readouts investigated by univariable modeling.__

Correlation of the 10 most influential factors affecting the combined mental health and quality of life scoring (__Figure \@ref(fig:fig-rf-components)__) with the anxiety (ANX) (__A__), depression (DPR) (__B__), overall mental health (OMH) (__C__) and quality of life (QoL) (__D__) rating was investigated by univariable, age- and sex-weighted Poisson regression (__Supplementary Table S5__). $\beta$ estimate values with 95% confidence intervals for the significant correlations in both the Austria/Tyrol (AT) and Italy/South Tyrol (IT) cohort are presented as forest plots.

prol.: prolonged, SIF: severe illness feeling, imp.: impaired, conc.: concentration, #: number, tired.day.: tiredness at day, pre-cov sleep disord.: sleep disorder before COVID-19, 3Q, 4Q: 3^rd^ and 4^th^ quartile, persist.: persistent.

\newpage

```{r fig-clustering, fig.width = mm_inch(paper_figures$clustering$w), fig.height = mm_inch(paper_figures$clustering$h), dpi = 600, fig.cap = 'Clustering of the study participants by the most influential factors affecting the combined mental health and quality of life scoring.'}

paper_figures$clustering$plot

```

__Figure \@ref(fig:fig-clustering). Clustering of the study participants by the most influential factors affecting the combined mental health and quality of life scoring.__

Study participants were assigned to the Low Risk (LR), Intermediate Risk (IR) and High Risk (HR) subsets by clustering analysis of the most influential factors affecting the combined mental health and quality of life scoring (__Figure \@ref(fig:fig-rf-components)__) by the self-organizing map (SOM, 11 $\times$ 11 hexagonal grid, Jaccard distance between participants) and the hierarchical clustering (Ward D2 method, Euclidean distance between the SOM nodes) algorithms. Presence/ absence of the features is presented as heat maps. N numbers of individuals assigned to the clusters are presented next to the plots.

prol.: prolonged, SIF: severe illness feeling, imp.: impaired, conc.: concentration, #: number, tired.day.: tiredness at day, pre-cov sleep disord.: sleep disorder before COVID-19, 3Q, 4Q: 3^rd^ and 4^th^ quartile, persist.: persistent.

\newpage

```{r fig-clust-mental, fig.width = mm_inch(paper_figures$clust_mental$w), fig.height = mm_inch(paper_figures$clust_mental$h), dpi = 600, fig.cap = 'Mental health scoring, depression and anxiety prevalence in the mental health risk clusters'}

paper_figures$clust_mental$plot

```

__Figure \@ref(fig:fig-clust-mental). Mental health scoring, depression and anxiety prevalence in the mental disorder risk clusters.__

Study participants were assigned to the Low Risk (LR), Intermediate Risk (IR) and High Risk (HR) subsets as presented in __Figure \@ref(fig:fig-clustering)__.

__(A)__ Rating of anxiety (ANX), depression (DPR), overall mental health (OMH) and quality of life (QoL) in the clusters presented as violin plots, diamonds with whiskers represent medians with IQRs. Statistical significance was assessed by Kruskal-Wallis test. 

__(B)__ Frequency of positive depression (DPR+) and anxiety (ANX+) screening in the clusters. Statistical significance was assessed by $\chi^2$ test.

P values corrected for multiple comparisons with Benjamini-Hochberg method are shown in the plots. N numbers of individuals assigned to the clusters are presented next to the plots.

\newpage

```{r pre-cov-da-cov-sympt, fig.width = mm_inch(paper_figures$pre_cov_da_cov_sympt$w), fig.height = mm_inch(paper_figures$pre_cov_da_cov_sympt$h), dpi = 600, fig.cap = 'Depression or anxiety before COVID-19 and COVID-19 symptom burden.'}

paper_figures$pre_cov_da_cov_sympt$plot

```

__Figure \@ref(fig:pre-cov-da-cov-sympt). Depression or anxiety before COVID-19 and COVID-19 symptom burden.__

Association of depression or anxiety before COVID-19 with the overall number of acute and persistent COVID-19 symptoms and neurocognitive (NC) COVID-19 symptoms was assessed by Mann-Whitney U test. Symptom numbers are presented as violin plots, diamonds with whiskers represent medians with IQRs. p values corrected for multiple comparisons with Benjamini-Hochberg method are shown in plot sub-headings. N numbers of observations are indicated below the plots.